Versatile Robust Clustering of Ad Hoc Cognitive Radio Network

### 2016-12-04 journal-draft/bare_jrnl_comsoc.tex
change the notation h_C--> h(C)

A distributed clustering scheme ROSS is proposed in~\cite{Li11_ROSS} under the game theoretic framework . 
Compared with the clustering schemes introduced above, the clusters are formed faster and the clusters possess more CCs within and among clusters than SOC.
But this work doesn't intervene the outcome of singleton clusters, although the cluster sizes are not as divergent as SOC.
Furthermore, this work doesn't consider the robustness of clusters against the increasing activity of primary users, which leaves their claim of robustness unverified.
%
This paper is on the basis of the work in~\cite{Li11_ROSS}, but extends in two directions.
First, this paper sticks to a new metric of robustness, and considers the clusters when primary users' activity becomes more dynamic.
Second, this paper proposes size control mechanism, which solves the problem of devergence of clusters sizes in ~\cite{Li11_ROSS} and~\cite{LIU_TMC11_2}.
Besides, this paper provides a comprehensive analysis of the robust clustering problem and proposes the centralized solution. 
%We stick to the motivation of forming robust clusters in CRN \ie let more CR users benefited from the cooperative decision making which is due to the clusters. 
%We propose both centralized and distributed clustering schemes, which result in more CR users in the clusters composed with multiple CR users, not only when the clusters are just formed, but also afterwards when the primary users' activity changes.
%Besides, both centralized and distributed schemes take the cluster size into consideration, and both can control the sizes of the formed clusters to certain extent.
%In particular, the decentralized clustering approach ROSS (RObust Spectrum Sharing) is able to form clusters with desired sizes.
%When compared with other distributed works, ROSS involves smaller signaling overhead and the generated clusters are significantly more robust against the primary users which appear after the clusters are formed.
%thesis
%and the generated clusters are more robust than the other clustering scheme which has claims on cluster robustness, \ie more secondary users residing in clusters against increasing influence from primary users.
%
%ROSS selects cluster head through coordination within its neighborhood, and then cluster membership is decided locally and its fast convergence is proved under game theoretic framework. 
%For our scheme we can prove convergence in cluster formation phase and resolve ambiguities with respect to cluster membership in a game-theoretic setting. 
The new extensions are made on basis of ROSS and its light weighted version, the latter involves less overheads thus are more suitable for the scenario where fast deployment is desired.
Throughout this paper, we refer to the clustering schemes on the basis of ROSS as \textit{variants of ROSS}, which include the fast versions and that with size control function.

@@

A cluster $C$ is a set of secondary nodes, which consists of a cluster head $h_C$ and a number of cluster members.
The cluster head is able to communicate with any cluster member directly.
$\text{Nb}(i)$ denotes node $i$'s neighborhood which consists of all its one hop neighbors.
%For any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.
Cluster size of $C$ is written as $|C|$.
Cluster $C(i)$ means the cluster head of this cluster is $i$.
$ K(C) = \cap_{i\in C} K_i$, $K(C)$, which denotes the set of CCs in cluster $C$.
%Clustering is performed periodically, because the secondary users are mobile and the primary users change their operation channels, thus the channel availability on secondary users changes accordingly.
The notations used in the system model and the following problem description are listed in Table~\ref{tab1}.
\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & set of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h_C$ & the cluster head of a cluster C\\
$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$f(C)$ & the number of CCs of a cluster $C$, which is used \\
& in the problem description\\
$C_i$  & the $i$th potential cluster in a collection of clusters \\
 $\mathcal{G}$ & a collection of some possible clusters in $\mathcal{N}$\\
 $\mathcal{S}$ & the collection of all the possible clusters in $\mathcal{N}$\\
\bottomrule
\end{tabular}
\end{table}

@@

As introduced in Section~\ref{intro}, in order to be robust against primary users' activity, the formed clusters should have more CCs to expect longer time expectancy.
Meanwhile, the sizes of the formed clusters should not diverge from the desired size greatly.
The formation of small clusters or the \textit{singleton clusters}, \ie the cluster which has only one CR node, contradicts the motivation of forming clusters, as the benefits brought in by the collective of the cluster members are compromised.
On the other hand, large clusters are not preferred in some scenarios neither, \eg for the CRN composed with resource limited users, managing the cluster members in a large cluster is a substantial burden. 
Hence, the cluster size should fall in a desired range according to different application scenarios~\cite{Chen04clusteringalgorithms, capacity_cluster_06}.
Considering the above mentioned requirements, we present the definition of robust clustering problem as follows.

@@

The procedure of the proof also illustrates the time needed to conduct Algorithm~\ref{clustering:theorem}. 
Consider an extreme scenario, where all the secondary nodes sequentially execute Algorithm~\ref{alg0}, \ie they constitute a list as discussed in the example in the proof.
%
If one step can be finished within certain time $T$, then the total time needed for the network to conduct Algorithm~\ref{clustering:theorem} is $N*T$.
In other scenarios, as Algorithm~\ref{alg0} can be executed concurrently by secondary users which locate in different places, the needed time can be considerably reduced.
%According to Theorem~\ref{clustering:theorem}, Phase I completes within a reasonable amount of time.
%
%This judgement is conducted periodically, and phase I ends after every node ascertains it is cluster head or not.
Let us apply Algorithm~\ref{alg0} to the example shown in Figure~\ref{fig1}.
Node $B$ and $H$ have the same individual connectivity degree, i.e., $d_B=d_H$. As $g_H=2>g_B=1$, node $H$ becomes the cluster head and cluster $C_H$ is $\{H, B, A, G\}$.

>>>>

A distributed clustering scheme ROSS is proposed in~\cite{Li11_ROSS} under the game theoretic framework. 
Compared with the clustering schemes introduced above, the clusters are formed faster and the clusters possess more CCs within and among clusters than SOC.
But this work doesn't intervene the outcome of singleton clusters, although the cluster sizes are not as divergent as SOC.
Furthermore, this work doesn't consider the robustness of clusters against the increasing activity of primary users, which leaves their claim of robustness unverified.
%
This paper is on the basis of the work in~\cite{Li11_ROSS}, but extends in two directions.
First, this paper sticks to a new metric of robustness, and considers the clusters when primary users' activity becomes more dynamic.
Second, this paper proposes size control mechanism, which solves the problem of devergence of clusters sizes in ~\cite{Li11_ROSS} and~\cite{LIU_TMC11_2}.
Besides, this paper provides a comprehensive analysis of the robust clustering problem and proposes the centralized solution. 
%We stick to the motivation of forming robust clusters in CRN \ie let more CR users benefited from the cooperative decision making which is due to the clusters. 
%We propose both centralized and distributed clustering schemes, which result in more CR users in the clusters composed with multiple CR users, not only when the clusters are just formed, but also afterwards when the primary users' activity changes.
%Besides, both centralized and distributed schemes take the cluster size into consideration, and both can control the sizes of the formed clusters to certain extent.
%In particular, the decentralized clustering approach ROSS (RObust Spectrum Sharing) is able to form clusters with desired sizes.
%When compared with other distributed works, ROSS involves smaller signaling overhead and the generated clusters are significantly more robust against the primary users which appear after the clusters are formed.
%thesis
%and the generated clusters are more robust than the other clustering scheme which has claims on cluster robustness, \ie more secondary users residing in clusters against increasing influence from primary users.
%
%ROSS selects cluster head through coordination within its neighborhood, and then cluster membership is decided locally and its fast convergence is proved under game theoretic framework. 
%For our scheme we can prove convergence in cluster formation phase and resolve ambiguities with respect to cluster membership in a game-theoretic setting. 
The new extensions are made on basis of ROSS and its light weighted version, the latter involves less overheads thus are more suitable for the scenario where fast deployment is desired.
Throughout this paper, we refer to the clustering schemes on the basis of ROSS as \textit{variants of ROSS}, which include the fast versions and that with size control function.

@@

A cluster $C$ is a set of secondary nodes, which consists of a cluster head $h(C)$ and a number of cluster members.
The cluster head is able to communicate with any cluster member directly.
$\text{Nb}(i)$ denotes node $i$'s neighborhood which consists of all its one hop neighbors.
%For any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.
Cluster size of $C$ is written as $|C|$.
Cluster $C(i)$ means the cluster head of this cluster is $i$.
$ K(C) = \cap_{i\in C} K_i$, $K(C)$, which denotes the set of CCs in cluster $C$.
%Clustering is performed periodically, because the secondary users are mobile and the primary users change their operation channels, thus the channel availability on secondary users changes accordingly.
The notations used in the system model and the following problem description are listed in Table~\ref{tab1}.
\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & set of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h(C)$ & the cluster head of a cluster C\\
$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$f(C)$ & the number of CCs of a cluster $C$, which is used \\
& in the problem description\\
$C_i$  & the $i$th potential cluster in a collection of clusters \\
 $\mathcal{G}$ & a collection of some possible clusters in $\mathcal{N}$\\
 $\mathcal{S}$ & the collection of all the possible clusters in $\mathcal{N}$\\
\bottomrule
\end{tabular}
\end{table}

@@

As introduced in Section~\ref{intro}, in order to be robust against primary users' activity, the formed clusters should have more CCs to expect longer life expectancy.
On the other hand, the sizes of the formed clusters should not diverge from the desired size greatly.
The formation of small clusters or the \textit{singleton clusters}, \ie the cluster which has only one CR node, contradicts the motivation of forming clusters, as the benefits brought in by the collective of the cluster members are compromised.
On the other hand, large clusters are not preferred in some scenarios neither, \eg for the CRN composed with resource limited users, managing the cluster members in a large cluster is a substantial burden. 
Hence, the cluster size should fall in a desired range according to different application scenarios~\cite{Chen04clusteringalgorithms, capacity_cluster_06}.
Considering the above mentioned requirements, we present the definition of robust clustering problem as follows.

@@

The procedure of the proof also illustrates the time needed to conduct Algorithm~\ref{clustering:theorem}. 
Consider an extreme scenario, where all the secondary nodes sequentially execute Algorithm~\ref{alg0}, \ie they constitute a list as discussed in the example in the proof.
%
If one step can be finished within certain time $T$, then the total time needed for the network to conduct Algorithm~\ref{clustering:theorem} is $N*T$.
In other scenarios, as Algorithm~\ref{alg0} can be executed concurrently by secondary users which locate in different places, the needed time can be considerably reduced.
%According to Theorem~\ref{clustering:theorem}, Phase I completes within a reasonable amount of time.
%
%This judgement is conducted periodically, and phase I ends after every node ascertains it is cluster head or not.
Let us apply Algorithm~\ref{alg0} to the example shown in Figure~\ref{fig1}.
Node $B$ and $H$ have the same individual connectivity degree, i.e., $d_B=d_H$. As $g_H=2>g_B=1$, node $H$ becomes the cluster head and cluster $C(H)$ is $\{H, B, A, G\}$.
